\pdftitle

### 2018-05-13 maxNt_v2.tex
V2

%% Author: PGL  Porta Mana
%% Created: 2015-11-04T09:20:08+0100
%% Last-Updated: 2018-05-12T15:29:50+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Report-no: ***
\documentclass{article}

@@

\usepackage[final,%
nonatbib]{nips_2017}

@@

\pdfmapline{+dummy-space <dummy-space.pfb}\pdfinterwordspaceon%

@@

A possible, precise definition is as follows. We want to predict, \ie\
assign a probability to, the population activity in a time bin, given the
population activities in the other bins:
\begin{equation}
  \label{eq:general_prediction}
  \p(\text{activity at one time bin} \|
  \text{activities at the remaining bins, and other info}).
\end{equation}
We can say that the simultaneous firing of at most two neurons
\enquote{explains} the activity if only the values of the first and second
empirical moments of the activities are relevant for the probability above:
\begin{multline}
  \label{eq:informal_sufficiency}
  \p(\text{activity at one time bin} \|
  \text{activities at the other bins, and other info}) ={}\\
  \p(\text{activity at some time bin} \|
  \text{1st \amp\ 2nd empirical moments, and other info});
\end{multline}
in statistical terms, the first and second moments are \emph{sufficient
  statistics}. Other definitions are possible, of course; the definition
above has the advantage of being mathematically very precise. In fact it
determines, from just the three basic laws of the probability calculus, the
mathematical form of the probability for the activity at any number of bins:
\begin{multline}
  \label{eq:sufficiency_maxent_family}
  \p(\text{activity at bins $1,2,\dotsc$} \|
  \text{other info}) ={}\\
  \shoveright{\int
    \pf(\text{activity at bin $1$} \|\ytheta)\,
    \pf(\text{activity at bin $2$} \|\ytheta)\dotsm
 \pf(\ytheta\| \text{other info})\,
  \di\ytheta,}
  \\[2\jot]
  \text{with \enquote{$\pf(\text{activity at time bin $i$} \|\ytheta)$}
    having maximum-entropy form}.
\end{multline}
The proof that the assumption of sufficient statistics leads to the
probability expression above is the Koopman-Pitman theorem
\cite{koopman1936,pitman1936,darmois1935} and can be found in
\citetext[\sect~4.5]{bernardoetal1994_r2000} and references therein. It can
be extended to higher moments and other statistics.

>>>>

%% Author: PGL  Porta Mana
%% Created: 2015-11-04T09:20:08+0100
%% Last-Updated: 2018-05-13T13:23:37+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Report-no: ***
\documentclass{article}

@@

\usepackage[final,%
nonatbib]{nips_2018}

@@

%\pdfmapline{+dummy-space <dummy-space.pfb}\pdfinterwordspaceon%

@@

A possible, precise definition is as follows. We want to predict, \ie\
assign a probability to, the population activity in a time bin, given the
population activities in the other bins:
\begin{equation}
  \label{eq:general_prediction}
  \p(\text{\small activity at one time bin} \|
  \text{\small activities at the remaining bins},\text{\small other info}).
\end{equation}
We can say that the simultaneous firing of at most two neurons
\enquote{explains} the activity if only the values of the first and second
empirical moments of the activities are relevant for the probability above:
\begin{multline}
  \label{eq:informal_sufficiency}
  \p(\text{\small activity at one time bin} \|
  \text{\small activities at the other bins},\text{\small other info}) ={}\\
  \p(\text{\small activity at some time bin} \|
  \text{\small 1st \amp\ 2nd empirical moments},\text{\small other info});
\end{multline}
in statistical terms, the first and second moments are \emph{sufficient
  statistics}. Other definitions are possible, of course; the definition
above has the advantage of being mathematically very precise. In fact it
determines, from just the three basic laws of the probability calculus, the
mathematical form of the probability for the activity at any number of bins:
\begin{equation}
  \label{eq:sufficiency_maxent_family}
  \begin{gathered}
  \p(\text{\small activity at bins $1,2,\dotsc$} \|
  \text{\small other info}) =
  \int\Bigl[\prod_i 
    \pf(\text{\small activity at bin } i \|\ytheta) \Bigr]\,
 \pf(\ytheta\| \text{\small other info})\,
  \di\ytheta,
  \\[\jot]
  \text{with \enquote{$\pf(\text{\small activity at time bin } i \|\ytheta)$}
    having maximum-entropy form with multipliers $\ytheta$};
\end{gathered}
\end{equation}
The proof that the assumption of sufficient statistics leads
to the probability expression above is the Koopman-Pitman theorem
\cite{koopman1936,pitman1936,darmois1935} and can be found in
\textcite[\sect~4.5]{bernardoetal1994_r2000} and references therein. It can
be extended to higher moments and other statistics.

When many data are available (and unfortunately also when they aren't), the
conditional density for the Lagrange multipliers,
$\pf(\ytheta\| \text{data, other info})$ is strongly peaked at the unique
values determined by the inverse problem \cite{meadetal1984}, and the
probability for the population activity is well-approximated by a
maximum-entropy distribution \cite{jaynes1957,sivia1996_r2006,meadetal1984}
with those Lagrange multipliers. This kind of distributions has been used
in neuroscience at least since the 1990s to study correlations from diverse
points of view, for diverse purposes, and with data recorded from diverse
brain areas
\cite{mackay1991,martignonetal1995,bohteetal2000,amarietal2003,schneidmanetal2006,shlensetal2006,mackeetal2009b,roudietal2009c,tkaciketal2009,gerwinnetal2009,mackeetal2011,mackeetal2011b,ganmoretal2011,granotatedgietal2013,tkaciketal2014b,moraetal2015,shimazakietal2015}.
Part of this literature focused on the range of applicability of maximum-entropy
distributions
\cite{tkaciketal2006,roudietal2009,roudietal2009b,barreiroetal2010b,mackeetal2013}.

This kind of studies would have little significance if we did not expect
their results to generalize beyond the particular recorded sample. In other
words, we implicitly or explicitly expect that the conclusions drawn with
maximum-entropy methods about the recorded neurons should somehow
generalize or extrapolate at least to whole areas those neurons are recorded from,
\eg\ retina or motor cortex. We are thus assuming that the recorded sample
is a \emph{representative sample} of that area.

Whether such extrapolation makes biological sense is a difficult question,
to be analysed case by case. The purpose

The point that we want to stress is that

From the observation of these samples we expect
to learn something about all other neurons in the same brain area. That is,
we assume that they are a \enquote{representative sample}.
